Modal Abstractions for Virtualizing Memory Addresses

### 2023-04-11 programlogic.tex
in the middle of refactoring

\todo[inline]{We should drop all talk about PDE, PDPT, etc., and just use L1-L4}
\paragraph{Physical-Memory points-to} Our physical memory points-to relation in Figure \ref{fig:physicalpointsto} for a page table entry (e.g. \textsf{PDE} in Figure \ref{fig:pagetables}) abstracts masking -- $|^{52}$ for frame for and $|^{12}$ for offset -- through logical nested maps.  
\begin{figure}[t]
\[
\begin{array}{cl}
\pfpointsto\locsf\vpts\qfrac\ppts \stackrel{{\triangle}}{:=} & \nfpointsto{\mask\locsf\ft\generalentry}{\mask\locsf\tw\generalentry}\vpts\qfrac\naddr
\end{array}
\]
\caption{Physical Points-to with Nested Masking}
  \label{fig:physicalpointsto}
\end{figure}
Given the definition of physical page-pointsto assertion and the root address of virtual-address space kept in the control register \lstinline{cr3} as shown in Figure \ref{fig:pagetable},  one can build the physical address-translation for a virtual address (e.g. \textsf{va}) via abstracting the L4-L1 table traversal as the following:
\begin{itemize}
  \item Page Map Level-4 Translation (L4): Performs the first level address translation to get the PML4 entry (PML4E) in Figure \ref{fig:pagetables} by using the masks (\textsf{va}$|^{52}$(\lstinline{cr3}) to get the starting address of the PML4 table (PLM4T) and using the offset mask \textsf{va}$|^{12}$(\lstinline{cr3}). It is worth noting that our offset indices are 12 bits as all physical memory pointsto assertions has to satisfy being \textsf{\textit{aligned}}, e.g. [47-39] plus 3 alignment bits in Figure \ref{fig:pagetables}
    \[ \begin{array}{l}
      \hbox{(\TirNameStyle{{L4translate}})} \quad
      \exists \entryf \ldotp \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qone\naddr
      \end{array}
      \]
  \item Page Directory Pointer Level Translation (L3): Performs the second level address translation to get the PDP entry (PDPE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PLM4E}) to get the starting address of the PDP table (PDPT) and using the offset mask \textsf{va}$|^{12}$(\textsf{PLM4E})
    \[\begin{array}{l}
    \hbox{(\TirNameStyle{{L3translate}})} \quad
    \exists \entrytr \ldotp \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qtwo\naddr
    \end{array}
    \]
  \item Page Directory Table Level Translation (L2): Performs the second level address translation to get the PD entry (PDE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PDPE}) to get the starting address of the PD table and using the offset mask \textsf{va}$|^{12}$(\textsf{PDPE})
    \[ \begin{array}{l}
            \hbox{(\TirNameStyle{{L2translate}})} \quad 
       \exists \entrytw \ldotp \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qthree\naddr
      \end{array}
      \]
  \item Page Table Level Translation (L1): Performs the second level address translation to get the PT entry (PTE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PDE}) to get the starting address of the Page table (PT) and using the offset mask \textsf{va}$|^{12}$(\textsf{PDE})
    \[ \begin{array}{l}
      \hbox{(\TirNameStyle{{L1translate}})} \quad
      \exists \entryo \ldotp \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfour\naddr
      \end{array}
      \]
  \item Page Address Level Translation: Performs the second level address translation to get the physical page address in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PTE}) to get the starting address of the physical page address table and using the offset mask \textsf{va}$|^{12}$(\textsf{PTE})
    \[ \begin{array}{l}
      \hbox{(\TirNameStyle{{PageLebelAccess}})} \quad
      \exists \vpage \ldotp \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
      \end{array}
      \]

@@

\begin{figure*}
\begin{lstlisting}[language=Coq]
 Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage$ :vProp $\Sigma$ := 
  $\exists_{\entryf \;, \entrytr \;, \entrytw \;,\entryo} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo($\rtv,\entryf,\entrytr,\entrytw,\entryo$) $\ast \entryo \mapsto_{p} \vpage$.
 Definition L4_L1_PointsTo (rtv maddr entry entry0 entry1 pte :word 64) : vProp $\Sigma$ :=
  ((l4M52 maddr rtv) \$\sim$\ (l4off maddr rtv) $\mapsto_{a}$ {q1} entry  $\ast$
  (l3M52 maddr entry) \$\sim$\ (l3off maddr entry)  $\mapsto_{a}$ {q2} entry0 $\ast$ 
  (l2M52 maddr entry0) \$\sim$\ (l2off maddr entry0) $\mapsto_{a}$ {q3} entry1 $\ast$
  (l1M52 maddr entry1) \$\sim$\ (l1off maddr entry1) $\mapsto_{a}$ {q4} pte ).
\end{lstlisting}
\caption{A Strong Virtual Points-to Relation}
\todo[inline,color=red]{Colin says: shouldn't that fraction be a q, not a 1, in the second line?}
  \label{fig:strongvirtualpointsto}
\end{figure*}

%then we reach to a physical page addr (e.g. \textsf{pa}) 
%\[  \paddr\mapsto_{\textsf{a}}\{1\}\vsome \ast \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfrac\naddr \]

>>>>

\todo[inline]{We should drop all talk about PDE, PDPT, etc., and just use L1-L4}
\paragraph{Physical-Memory points-to} Our physical memory points-to relation in Figure \ref{fig:physicalpointsto} for a page table entry (e.g. \textsf{PDE} in Figure \ref{fig:pagetables}) abstracts masking -- $|^{52}$ for frame for and $|^{12}$ for offset -- through logical nested maps.

Given the definition of physical page-pointsto assertion and the root address of virtual-address space kept in the control register \lstinline{cr3} as shown in Figure \ref{fig:pagetable},  one can build the physical address-translation for a virtual address (e.g. \textsf{va}) via abstracting the L4-L1 table traversal as the following:
\begin{itemize}
  \item Page Map Level-4 Translation (L4): Performs the first level address translation to get the PML4 entry (PML4E) in Figure \ref{fig:pagetables} by using the masks (\textsf{va}$|^{52}$(\lstinline{cr3}) to get the starting address of the PML4 table (PLM4T) and using the offset mask \textsf{va}$|^{12}$(\lstinline{cr3}). It is worth noting that our offset indices are 12 bits as all physical memory pointsto assertions has to satisfy being \textsf{\textit{aligned}}, e.g. [47-39] plus 3 alignment bits in Figure \ref{fig:pagetables}
    \[ \begin{array}{l}
      \hbox{(\TirNameStyle{{L4translate}})} \quad
      \exists \entryf \ldotp \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qone\naddr
      \end{array}
      \]
  \item Page Directory Pointer Level Translation (L3): Performs the second level address translation to get the PDP entry (PDPE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PLM4E}) to get the starting address of the PDP table (PDPT) and using the offset mask \textsf{va}$|^{12}$(\textsf{PLM4E})
    \[\begin{array}{l}
    \hbox{(\TirNameStyle{{L3translate}})} \quad
    \exists \entrytr \ldotp \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qtwo\naddr
    \end{array}
    \]
  \item Page Directory Table Level Translation (L2): Performs the second level address translation to get the PD entry (PDE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PDPE}) to get the starting address of the PD table and using the offset mask \textsf{va}$|^{12}$(\textsf{PDPE})
    \[ \begin{array}{l}
            \hbox{(\TirNameStyle{{L2translate}})} \quad 
       \exists \entrytw \ldotp \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qthree\naddr
      \end{array}
      \]
  \item Page Table Level Translation (L1): Performs the second level address translation to get the PT entry (PTE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PDE}) to get the starting address of the Page table (PT) and using the offset mask \textsf{va}$|^{12}$(\textsf{PDE})
    \[ \begin{array}{l}
      \hbox{(\TirNameStyle{{L1translate}})} \quad
      \exists \entryo \ldotp \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfour\naddr
      \end{array}
      \]
  \item Page Address Level Translation: Performs the second level address translation to get the physical page address in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PTE}) to get the starting address of the physical page address table and using the offset mask \textsf{va}$|^{12}$(\textsf{PTE})
    \[ \begin{array}{l}
      \hbox{(\TirNameStyle{{PageLebelAccess}})} \quad
      \exists \vpage \ldotp \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
      \end{array}
      \]

@@

\begin{figure*}
  \begin{lstlisting}[language=Coq]
 Definition $\pfpointsto\locsf\vpts\qfrac\ppts$ :vProp $\Sigma$ := pmapsto_def_v (pfn w) (page_off w) q v.
 Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage$ :vProp $\Sigma$ := 
  $\exists_{\entryf \;, \entrytr \;, \entrytw \;,\entryo} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo($\rtv,\entryf,\entrytr,\entrytw,\entryo$) $\ast$ addr_L1$(\vaddr,\entryo) \mapsto_{p} \vpage$.
 Definition L4_L1_PointsTo (rtv maddr entry entry0 entry1 pte :word 64) : vProp $\Sigma$ :=
  ((l4M52 maddr rtv) \$\sim$\ (l4off maddr rtv) $\mapsto_{a}$ {q1} entry  $\ast$
  (l3M52 maddr entry) \$\sim$\ (l3off maddr entry)  $\mapsto_{a}$ {q2} entry0 $\ast$ 
  (l2M52 maddr entry0) \$\sim$\ (l2off maddr entry0) $\mapsto_{a}$ {q3} entry1 $\ast$
  (l1M52 maddr entry1) \$\sim$\ (l1off maddr entry1) $\mapsto_{a}$ {q4} pte ).
\end{lstlisting}
\caption{A Strong Virtual Points-to Relation}
\todo[inline,color=red]{Colin says: shouldn't that fraction be a q, not a 1, in the second line?}
  \label{fig:strongvirtualpointsto}
\end{figure*}
